Think deep for reef resilience

### 2025-07-16 paper/paper_draft.qmd
Add stub section for depth to fill in later
Possible discussion text on connectivity modelling methods
Still need to emphasise it's importance. And would be good to add similar changes for DHW-tolerance and possible improvements in understanding

@@

ADRIA-CoralBlox was run for 4096 **sampled** scenarios per GCM. A median absolute coral
cover timeseries for each reef was calculated for each GCM. For timeseries clustering reef
cover timeseries were scaled by the total area of the reef to represent timeseries as a
percentage of the total reef area. The proportion of each functional group occupying the
reef (group evenness) was also investigated as a metric for timeseries analysis, but was
found to not sufficiently capture the declined condition of reefs suffering from very low
absolute coral cover.

@@

It has been suggested that a primary facet of reef resilience is the degree of larval
connectivity between reefs and the supply of recruits that a reef receives following
disturbance [@hock2017]. It was expected that resilient reefs would have greater incoming
than outgoing larval connectivity to promote recovery in years following thermal stress,
while non-resilient reefs would have greater outgoing than incoming connectivity inhibiting
recruitment of corals ability post thermal stress. In contrast, our results suggest that
there are few differences between resilient and non-resilient reefs in connectivity metrics
such as the ratio of outgoing to incoming larval connectivity and total connectivity
strength. Hock et al. [-@hock2017] used larval connectivity modelling to identify resilient
reefs that may supply up to 47% of the wider Great Barrier Reef ecosystem with larvae
following disturbance. A critique of this approach [@bode2018] is that results obtained are
highly sensitive to the biophysical dispersal model used, limiting the application of
specific identified reefs. Reef depth was not considered in these investigations of reef
network resilience. These investigations were performed using prior generations of larval
connectivity models parameterised for 2008-2013 [@hock2017] and 1996-2002 [@bode2018] time
periods respectively. Using updated 1km resolution larval connectivity across the whole GBR,
our results suggest that larval connectivity alone is not as important for reef resilience
and recovery as environmental refugia from thermal stress such as depth.

>>>>

### Depth estimation

\*\* Describe process of estimating median depth for each reef \*\*

@@

ADRIA-CoralBlox was run for 4096 **sampled** scenarios for each of the five GCMs.
A median absolute coral cover timeseries for each reef was calculated for each GCM.
For timeseries clustering reef cover timeseries were scaled by the total area of the reef
to represent timeseries as a percentage of the total reef area. The proportion of each
functional group occupying the reef (group evenness) was also investigated as a metric for
timeseries analysis, but was found to not sufficiently capture the declined condition of reefs
suffering from very low absolute coral cover.

@@

It is a strongly held conclusion in coral reef ecology that a primary facet of reef resilience 
is the degree of larval connectivity between reefs and the supply of recruits that a 
reef receives following disturbance [@hock2017] **expand citations**. It was expected that resilient reefs would
have greater incoming than outgoing larval connectivity to promote recovery in years 
following thermal stress, while non-resilient reefs would have greater outgoing than 
incoming connectivity inhibiting recruitment of corals ability post thermal stress. In 
contrast, our results suggest that there are few differences between resilient and 
non-resilient reefs in connectivity metrics such as the ratio of outgoing to incoming larval 
connectivity and total connectivity strength. Hock et al. [-@hock2017] used larval 
connectivity modelling to identify resilient reefs that may supply up to 47% of the wider 
Great Barrier Reef ecosystem with larvae following disturbance. A critique of this 
approach [@bode2018] is that results obtained are highly sensitive to the biophysical 
dispersal model used, limiting the application of specific identified reefs. Reef depth was
not considered in these investigations of reef network resilience. These investigations 
were performed using prior generations of larval connectivity models parameterised for 
2008-2013 [@hock2017] and 1996-2002 [@bode2018] time periods respectively. Using updated 
1km resolution larval connectivity across the whole GBR, with a limited temporal sample of
particle flows, our results suggest that larval connectivity alone is not as important for
reef resilience and recovery as environmental refugia from thermal stress such as depth. 
Our findings may differ due to the highly uncertain nature of connectivity methods used 
**citation**, the way larval flows and juvenile recruitment are implemented in ADRIA - 
CoralBlox, and the magnitude of early bleaching events decreasing the breeding capacity 
of strongly connected reefs. Further improvements in connectivity modelling are necessary 
to provide a robust comparison of the factors influencing reef resilience.
